Hierarchically Supervised Latent Dirichlet Allocation

### 2011-06-02 paper/abstract.tex
subfigure intro (not finished)

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
model for hierarchically and multi-labeled bag-of-word data.  Examples of
such data include Web pages and their placement in directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data is also of interest.
%We define a model that uses probit regressors on a conditionally dependent
%label hierarchy tied to latent Dirichlet allocation (LDA). 
We demonstrate HSLDA on large-scale data from clinical document labeling and
retail product categorization tasks. We show that leveraging the structure from
hierarchical labels improves out-of-sample label prediction substantially when
compared to models that do not. %Furthermore, we show evidence
%that the resulting HSLDA topics are more descriptive of the underlying data
%than sLDA topics, which ignore the label hierarchy.

>>>>

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
model for hierarchically and multiply labeled bag-of-word data.  Examples of
such data include web pages and their placement in directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data is also of interest.
%We define a model that uses probit regressors on a conditionally dependent
%label hierarchy tied to latent Dirichlet allocation (LDA). 
We demonstrate HSLDA on large-scale data from clinical document labeling and
retail product categorization tasks. We show that leveraging the structure from
hierarchical labels improves out-of-sample label prediction substantially when
compared to models that do not. %Furthermore, we show evidence
%that the resulting HSLDA topics are more descriptive of the underlying data
%than sLDA topics, which ignore the label hierarchy.

### 2011-06-02 paper/experiments.tex
subfigure intro (not finished)

Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
relative to the three comparison models. In figure %\ref{fig:1b}.

%\begin{figure}%[htbp]
%\begin{center}
%%\subfloat[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
%\subfigure[][]{\label{1b}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
%%\subfloat[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
%\subfigure[][]{\label{1d}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
%\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records 
%(\subref{} and \subref{1b}). Out-of-sample Amazon product category predictions from 
%product free-text descriptions (\subref{} and \subref{1d}). In all figures, solid is 
%HSLDA, dashed are independent regressors + sLDA (hierarchical 
%constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
%tree-conditional regressions. \subref{} and \subref{} show predictive performance as a function of the 
%auxiliary variable threshold and \subref{1b} and \subref{1d} show predictive perfmance as a function of the
%prior mean on regression coefficients.}
%\label{fig:main_results}
%\end{center}
%\end{figure}

>>>>

Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
relative to the three comparison models. In figure \ref{fig:1b}.

\begin{figure}%[htbp]
\begin{center}
%\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
%\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
\subfigure[][]{\label{fig:1d}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records 
(\subref{1a} and \subref{fig:1b}). Out-of-sample Amazon product category predictions from 
product free-text descriptions (\subref{1c} and \subref{fig:1d}). In all figures, solid is 
HSLDA, dashed are independent regressors + sLDA (hierarchical 
constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
tree-conditional regressions. \subref{} and \subref{} show predictive performance as a function of the 
auxiliary variable threshold and \subref{fig:1b} and \subref{fig:1d} show predictive perfmance as a function of the
prior mean on regression coefficients.}
\label{fig:main_results}
\end{center}
\end{figure}

### 2011-06-02 paper/introduction.tex
subfigure intro (not finished)

The task of multi-label classification, selecting the k-best labels for a given
instance, has been a topic of research for several years. One simplistic way to
carry out the classification is through a series of independent binary
classifiers, but this ignores the many inherent dependencies among the
labels. Thus, much work has been devoted on incorporating the co-occurence
patterns of the labels into the classification task. In this paper, we focus on
multi-label classification, where the labels are organized in a hierarchical
structure. Scenarios of use include, but are not limited to, placing webpages
into manually curated Internet directories~\cite{DMOZ}, categorizing images
according to a taxonomy, tagging product descriptions with catalogue
information~\cite{SNAP}, and assigning diagnosis codes to clinical
records~\cite{Challenge07}.

There are several challenges entailed in incorporating the hierarchical nature
of labels into the classification task. One pertains to the labeling itself: in
the datasets (especially real-world, noisy ones), for a given label, instances
labeled with it contribute positive instance, but it is unclear how to
determine the negative instances. In particular, how to treat the parent labels
of the selected ones?

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. In particular, we
extend supervised latent Dirichlet allocation (sLDA)~\cite{BleiMcAuliffe2008}
to take advantage of hierarchical supervision. We hypothesize that hierarchical
label information provides more information about labeling than considering
labels as a flat list.

We demonstrate our model on large, real-world datasets in the clinical and web
retail domains. We observe that hierarchical information is valuable when
incorporated into the learning and improves our primary goal of multi-label
classification. Following the trend observed in supervised topic modeling, we
note that the learned topic models are more representative of the underlying
data in both of our datasets~\cite{BleiMcAuliffe2008}.

@@

>>>>

There exist many sources of unstructured data that have been partially or
completely categorized by human editors.  In this paper we focus on
unstructured textual data that has been, at least in part, manually
categorized.  Examples include but are not limited to webpages and curated
hierarchical directories of the same \citep{DMOZ}, product descriptions and
catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes
(e.g.~hospital discharge summaries with 
International Classification of Disease 9th Revision, Clinical Modification
(ICD-9-CM) codes assigned \cite{ICD9}).  In this work we show how to combine
these two sources of information using a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. The models and techniques that we develop in
this paper are applicable in other domains as well, namely, any unstructured
representations of data that have been hierarchically classified (e.g.,~image
catalogs with bag-of-feature image representations).

There are several challenges entailed in incorporating a hierarchy of labels.
Given a large set of potential labels (often thousands), each instance has only
a small number of labels associated to it. There are no negative labeling
present in the data naturally, and the absence of a label cannot always be
interpreted as a negative labeling.

%The task of multi-label classification, selecting the k-best labels for a given
%instance, has been a topic of research for several years. One simplistic way to
%carry out the classification is through a series of independent binary
%classifiers, but this ignores the many inherent dependencies among the
%labels. Thus, much work has been devoted on incorporating the co-occurence
%patterns of the labels into the classification task.

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. In particular, we
extend supervised latent Dirichlet allocation (sLDA)~\cite{BleiMcAuliffe2008}
to take advantage of hierarchical supervision. We hypothesize that hierarchical
label information provides more information about labeling than considering
unstructured labels.

We demonstrate our model on large, real-world datasets in the clinical and web
retail domains. We observe that hierarchical information is valuable when
incorporated into the learning and improves our primary goal of multi-label
classification. %in two ways

@@

%In this work we extend supervised latent Dirichlet allocation (sLDA)
%\cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.  sLDA
%is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per
%document ``supervision'';  often taking the form of a single numerical or
%categorical ``label.''  More generally this supervision is just extra per
%document data;  for instance its quality or relevance (e.g.~online review
%scores), marks given to written work (e.g.~essay grades), or the number of
%times a web page is linked.  These labels are usually generatively modeled as
%having been conditionally drawn from some distribution that depends on the
%document-specific topic mixture.  It has been demonstrated that the signal
%provided by such supervision can result in better, task-specific document
%models and can also lead to good label prediction for out-of-sample data
%\cite{BleiMcAuliffe2008}.
